untitled

### 2016-04-23 model.tex
update

\begin{align}
  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0,\\
  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
\end{align}  
where $\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
square of the epicylic frequency. We assume $\kappa^2>0$. 
These criteria are applicable to any adiabatic fluid when $s$
is the usual definition of entropy. For the dusty problem, we may take 
$s=\ln{(1-\epsilon)}+\mathrm{const.}$ from the definition of $s$ and  
the equation of state (Eq. \ref{eos}) with uniform sound-speed.

@@

\begin{align} 
  &1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1+
    \tepsilon\right)^2}\frac{\p\ln{\rhog}}{\p\ln{r}}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}} +
  \tepsilon\frac{z^2}{\widetilde{H}^2}>0,   \\ 
&1 - \frac{\tepsilon
  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right) > 0,
\end{align}

\subsubsection{Order-of-magnitude estimates}
Let us assume the gas and dust distributions have a Gaussian
distribution vertically,
\begin{align}
  \rho_\mathrm{g,d} = \rho_\mathrm{g0,d0} \exp
      {\left(-\frac{z^2}{2H^2_\mathrm{g,d}}\right)}.
\end{align}
Then the dust-to-gas ratio is 
\begin{align}
  \tepsilon = \tepsilon_0
  \exp{\left(-\frac{z^2}{2\widetilde{H}^2}\right)}, 
\end{align}
with $\tepsilon_0= \rho_\mathrm{d,0}/\rho_{g_0}$, and  
\begin{align}
  \frac{1}{\widetilde{H}^2} \equiv \frac{1}{H_\mathrm{d}^2} -
  \frac{1}{H_\mathrm{g}^2}. 
\end{align}
We assume the dust layer is much thinner than the gas scale-height,
$H_\mathrm{d}\ll H_\mathrm{g}$, so that 
\begin{align}
 \widetilde{H} \sim H_\mathrm{d}. 
\end{align}

We further assume the gas and dust
distributions are smoothly varying in the radial direction so that  
\begin{align}
  \frac{\p}{\p r}\sim \frac{1}{r}, 
\end{align}
the disk is roughly Keplerian so that
\begin{align}
  \kappa \sim \OmK, 
\end{align}
and the gas sound-speed is 
\begin{align}
  c_s \sim h_\mathrm{g}r\OmK, 
\end{align}
where $h_\mathrm{g}\ll1$ is the aspect-ratio of the gas component.

Inserting the above approximations into Eq. \ref{criterion2}, and
evaluating it at $z = H_\mathrm{d}$, we find 
\begin{align}
1 + \frac{h_\mathrm{g}^2\tepsilon}{\left(1 + \tepsilon\right)^2}
\left[1 - \left(\frac{H_\mathrm{d}}{H_\mathrm{g}}\right)^2 +
  \frac{\tepsilon}{1+\tepsilon}\left(1 -
  \frac{H_\mathrm{d}}{H_\mathrm{g}}\right)   
\right]>0.
\end{align}

>>>>

\begin{align}
  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0,\\
  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
\end{align}
with $s = \ln{P/\rho^\gamma}$. These criteria are applicable to any adiabatic fluid. 
For the present dusty problem, we may take 
where $\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
square of the epicylic frequency. We assume $\kappa^2>0$. 
These criteria are applicable to any adiabatic fluid when $s$
is the usual definition of entropy. For the dusty problem, we may take 
$s=\ln{(1-\epsilon)}+\mathrm{const.}$ from the definition of $s$ and  
the equation of state (Eq. \ref{eos}) with uniform sound-speed.

@@

\subsubsection{Order-of-magnitude estimates}
Let us assume the gas and dust distributions have a Gaussian
distribution vertically,
\begin{align}
  \rho_\mathrm{g,d}(r,z) = \rho_\mathrm{g0,d0}(r) \exp
      {\left[-\frac{z^2}{2H^2_\mathrm{g,d}(r)}\right]}.
\end{align}
Then the dust-to-gas ratio is 
\begin{align}
  \tepsilon = \tepsilon_0(r)
  \exp{\left[-\frac{z^2}{2\widetilde{H}^2(r)}\right]}, 
\end{align}
with $\tepsilon_0= \rho_\mathrm{d,0}/\rho_{g_0}$, and  
\begin{align}
  \frac{1}{\widetilde{H}^2} \equiv \frac{1}{H_\mathrm{d}^2} -
  \frac{1}{H_\mathrm{g}^2}. 
\end{align}
%We assume the dust layer is much thinner than the gas scale-height,
%$H_\mathrm{d}\ll H_\mathrm{g}$, so that 
%\begin{align}
% \widetilde{H} \sim H_\mathrm{d}. 
%\end{align}

We further assume
%the gas and dust
%distributions are smoothly varying in the radial direction so that  
%\begin{align}
%  \frac{\p}{\p r}\sim \frac{p}{r}, 
%\end{align}
the disk is roughly Keplerian so that
\begin{align}
  \kappa \sim \OmK, 
\end{align}
and the gas sound-speed is 
\begin{align}
  c_s \sim h_\mathrm{g}r\OmK, 
\end{align}
where $h_\mathrm{g}\ll1$ is the aspect-ratio of the gas component.

%Inserting the above approximations into Eq. \ref{criterion2} and writing the result in terms of the
%dust-to-gas ratio $\tepsilon$,           
%we find 
%\begin{align}\label{ineq_est}
%  1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1 + \tepsilon\right)^2}
%  \left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right| 
%  \left[
%    \left|\frac{\p\ln{\rhog}}{\p\ln{r}}\right| - \frac{\widetilde{H}^2}{H^2_\mathrm{g}}\left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right|  
%    \right]>0
%\end{align}
%is required \emph{for stability.} 
%This inequality is easily satisifed in typical disk models where 
%the dust layer is much thinner than the gas scale-height, so
%that $\widetilde{H}\sim H_\mathrm{d}\ll H_\mathrm{g}$, and where the radial profiles smoothly vary on a scale of $r$.
%Then the magnitude the only destabilizing term (the third term) is negligible. According to Eq. \ref{ineq_est}, instability is on%ly possible if there is a rapid variation the radial profile of the dust-to-gas ratio.

\begin{align} 
  &1 + \frac{\tepsilon}{\left(1+
    \tepsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}} + 
  \frac{z^2}{\widetilde{H}^2}\right)>0,   \\ 
&1 - \frac{\tepsilon
  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right) > 0,
\end{align}
